Comparing R packages for calculation of persistent homology

### 2020-03-19 tdabench-paper/tdabench.Rmd
fixed phrasing

Based on the theoretical complexity and our results, alpha complexes are superior for point clouds with 3 or fewer dimensions.
This advantage becomes especially clear at high number of points.
However, without sufficient computational resources, alpha complexes were not usable for point cloud dimensions greater than 3.
If a point cloud has more than 3 dimensions, then it could undergo pre-processing with dimension reduction before using alpha complexes.
However, if data dimension cannot be reduced without significant information loss, then Vietoris-Rips complexes should be used.
Among the tested libraries, Ripser (wrapped by \CRANpkg{TDAstats}) has the fastest calculation time. 
GUDHI and Dionysus (wrapped by \CRANpkg{TDA}) significantly fall behind as feature dimension and number of points increase. 
GUDHI and Dionysus libraries, however, do have the option of returning the boundary matrix, which Ripser does not do as one of its time saving features.

While Vietoris-Rips complexes can handle high dimensional data well, it still significantly slows down when looking for higher dimensional topological features. 
The size of the boundary matrix rapidly increases and saturates the RAM even for small numbers of points (e.g. ~100 points when looking for 4 dimensional features). 
Looking for high dimensional features in a large high dimensional point cloud remains a challenge. 
Other persistent homology complex methods exist such as the Delaunay complex, Witness complex, etc. but they are not currently implemented in R. 
Future challenges would be creating and implementing algorithms that reduce the computational complexity of higher dimensional topological feature calculations for R.

>>>>

Based on the theoretical complexity and our results, alpha complexes are superior for point clouds with 3 or fewer dimensions.
This advantage becomes especially clear at high number of points.
However, without sufficient computational resources, alpha complexes were not usable for point cloud dimensions greater than 3.
If a point cloud has more than 3 dimensions, then it could undergo pre-processing with dimension reduction before using alpha complexes.
However, if data dimension cannot be reduced without significant information loss, then Vietoris-Rips complexes should be used.
Among the tested libraries, Ripser (wrapped by \CRANpkg{TDAstats}) has the fastest calculation time. 
GUDHI and Dionysus (wrapped by \CRANpkg{TDA}) significantly fall behind as feature dimension and number of points increase. 
GUDHI and Dionysus, however, do have the option of returning the boundary matrix, which Ripser does not store for the sake of optimization.

While Vietoris-Rips complexes can handle high-dimensional data well, calculation still significantly slows down when looking for higher dimension features. 
Boundary matrix size rapidly increases and can overwhelm available random access memory even for small numbers of points (e.g. 100 points when looking for 4 dimensional features). 
Looking for high-dimensional features in a large, high-dimensional point clouds remains a challenge. 
Other persistent homology complex methods exist such as the Delaunay complex, Witness complex, etc. but they are not currently implemented in R. 
Future challenges would be creating and implementing algorithms that reduce the computational complexity of higher dimensional topological feature calculations for R.
